---
title: "Week 4 - Probability"
author: "Matthew E. Aiello-Lammens"
date: "9/27/2021"
output: html_document
---

## Learning Objectives

* Identify and practice the uses of **marginal**, **joint**, and **conditional** probabilities in understanding issues of public health and health care coverage.

## Review of some Quiz 3 questions

### Based on question 2.10 in textbook

Consider the following data on health coverage and health status, as reported by the Behavioral Risk Factor Surveillance System. 


|                       | Excellent health| Very good health| Good health| Fair health| Poor health| Total |
|:----------------------|:---------:|:---------:|:----:|:----:|:----:|:-----:|
|**No health coverage** |459        |727        |854   |385   |99    |2524   |
|**Yes health coverage**|4198       |6245       |4821  |1634  |578   |17476  |
|**Total**              |4657       |6972       |5675  |2019  |677   |20000  |

1. If we draw a survey participant at random, what is the probability that they have excellent health **and** do not have health coverage?

This is an example of calculating a **joint probability**. Look for the cell that corresponds to the number of individuals who have **both** excellent health and no health coverage, then divide this number by the total number of survey participants.

```{r}
# define total number of survey participants
total_participants = 20000

# define number of participants that are in excellent health and do not have health coverage
exhealth_nocoverage = 459

# calculate percentage of participants with excellent health but no coverage
prob_exhealth_nocoverage = exhealth_nocoverage/total_participants

# print this value to screen
prob_exhealth_nocoverage
```

2. If we draw a survey participant at random, what is the probability that they do not have health coverage?

This is an example of calculating a **marginal probability**. Look for the *marginal cell* that corresponds to the *total number of participants with no health coverage*, then divide this number by the total number of survey participants.

```{r}
# define total number of survey participants reporting no health coverage
tot_nocoverage = 2524

# calculate percentage of participants reporting no health coverage
prob_nocoverage = tot_nocoverage / total_participants

# print this value to the screen
prob_nocoverage
```


3. TRUE or FALSE - The probability that a survey participant has poor health and has health coverage is greater than a participant has poor health and does not have health coverage.

To answer this question, we need to calculate two different joint probability: P(poor health AND health coverage) and P(poor health AND no health coverage). 

```{r}
prob_poorhealth_yescoverage = 578 / total_participants
prob_poorhealth_nocoverage = 99 / total_participants

prob_poorhealth_yescoverage
prob_poorhealth_nocoverage

```


**Alternative question**

TRUE or FALSE - The probability that a survey participant has poor health **given** they have health coverage is greater than the probability a participant has poor health **given** they do not have health coverage.

To answer this question, we need to calculate two different conditional probabilities: P(poor health | health coverage)  and P(poor health | no health coverage). To do this, we'll need to focus on each of the health coverage rows separately.

```{r}
# define total number of survey participants reporting no health coverage
tot_nocoverage = 2524

# define the number of participants with poor health and no health coverage
poorhealth_nocoverage = 99

# define total number of participants reporting poor health
tot_yescoverage = 17476

# define the number of participants with poor health and health coverage
poorhealth_yescoverage = 578

# Calculate the percentages (probabilities) for poor health values within each row separately

# P(poor health | health coverage) 
poorhealth_yescoverage / tot_yescoverage

# P(poor health | no health coverage) 
poorhealth_nocoverage / tot_nocoverage

```


4. If we draw a survey participant at random and determine they have poor health, what is the probability that they do not have health coverage? 

This is another conditional probability question. In this case, we are *given* the fact that the survey participant has poor health. So, let's only focus on the poor health column. Considering these numbers, what is the probability that the participant does not have health coverage?

```{r}
# define the total number of participants in poor health
tot_poorhealth = 677

# define the number of participants with poor health and no health coverage
poorhealth_nocoverage = 99

# define the number of participants with poor health and health coverage
poorhealth_yescoverage = 578

# calculate the percentages (probabilities) of each of the health coverage categories, given poor health

# P(no health coverage | poor health)
poorhealth_nocoverage / tot_poorhealth

# P(yes health coverage | poor health)
poorhealth_yescoverage / tot_poorhealth

```

We could have also applied the equation for conditional probability to the last two problems. 

Recall:

$$
P(A|B) = \frac{P(A \cap B)}{P(B)}
$$

Applying this to the above we would have:

$$
P(\text{no health coverage} | \text{poor health}) = \frac{P(\text{no health coverage AND poor health})} {P(\text{poor health})}
$$

Calculate probability of poor health

```{r}
# define the total participants with poor health
tot_poorhealth = 677

# P(poor health)
prob_poorhealth = tot_poorhealth / total_participants
```

Calculate probability of poor health AND no health coverage

```{r}
prob_poorhealth_nocoverage = poorhealth_nocoverage / total_participants
```

Then we can calculate $P(\text{no health coverage} | \text{poor health})$ as:

```{r}
prob_poorhealth_nocoverage / prob_poorhealth
```




