---
title: "Week 2 - Pre-meeting setup to run `learnr` tutorials"
author: "Matthew E. Aiello-Lammens"
date: "8/28/2020"
output: html_document
---

## Running `learnr` tutorials

**Before each synchronous small group meeting** you are expected to complete a self-paced tutorial in R/RStudio.
I am making these tutorial using tools in the package `learnr`. 
In order for you run these tutorials you will need to take a few initial setup steps in R, detailed here.
In these steps, we are installing what are called `packages` for R.
`packages` are collections of R code written to be 'add ons' to the base R application.

**For Windows users**

You will also need to download **RTools**. You can find this application at 
[https://cran.rstudio.com/bin/windows/RTools/](https://cran.rstudio.com/bin/windows/RTools/).
When you have installed this program **RESTART RStudio**.


```{r learnr_setup, eval=FALSE}
install.packages("learnr")
install.packages("devtools")
install.packages("tidyverse")

devtools::install_github("rstudio-education/gradethis")
devtools::install_github("rundel/learnrhash")
devtools::install_github("mlammens/MAT141learnrMods")

```

Once this code is all set-up, we can initiate a tutorial using the following commands.

First, we can see all of the `learnr` tutorials available in a package:

```{r}
learnr::run_tutorial(package = "MAT141learnrMods")
```

Next, we can select the tutorial we want to run. **Note my use of quotes!**

```{r, eval=FALSE}
learnr::run_tutorial(name = "Short_Intro_to_R", package = "MAT141learnrMods")
```

The tutorial module should start in a new web browser tab.


In order to keep up with my updates to the package containing the `learnr` tutorials, you will need to run the following code after I post an announcement about an updated or added tutorial.

```{r learnr_update, eval=FALSE}
devtools::install_github("mlammens/MAT141learnrMods")
```
