---
title: "Week 8 - Class Plan - Hypothesis Testing - Fill in the Blanks"
author: "Matthew Aiello-Lammens"
date: "10/27/2021"
output:
  html_document:
    toc: yes
---

# Worked example on 1-sample hypothesis test

## *More* work with the Youth Risk Behavior Surveillance System (YRBSS) data

This week, we are again going to use the Youth Risk Behavior Surveillance System (YRBSS) data set.
Let's begin by loading the data into our environment.

```{r}
library(openintro)
data(yrbss)
```

Next, draw a random sample of these data, with sample size equal to 100.

```{r}
set.seed(1981)
yrbss_samp <- yrbss[sample(x = 1:nrow(yrbss), replace = FALSE, size = 100), ]
```


## Confidence Intervals

Calculate the **sample mean** and the **95% Confidence Interval** for **height** based on your sample. Recall that the **Confidence Interval** is:

$$
\bar{x} \pm 1.96\times SE \\
$$

or, otherwise written as:

$$
\bar{x} \pm 1.96\times \frac{s}{\sqrt{n}}
$$


```{r}

```

## Compare to population mean

Calculate the **population mean** for **height** for the `yrbss` data set. 
Does our **95% confidence interval** calculated in above *include* the true population mean?


```{r}

```



## Z-stat

Using the **population mean** calculated above, and the sample mean calculated even further above, calculate the $z$ statistic associated with your sample mean. Is your sample mean significantly different from the population mean?


```{r}


```





