---
title: "Week 8 - Class Plan - Hypothesis Testing - Fill in the Blanks"
author: "Matthew Aiello-Lammens"
output:
  html_document:
    toc: yes
---

# Worked example on 1-sample hypothesis test

## *More* work with the Youth Risk Behavior Surveillance System (YRBSS) data

This week, we are again going to use the Youth Risk Behavior Surveillance System (YRBSS) data set.
Let's begin by loading the data into our environment.

```{r}
library(openintro)
data(yrbss)
```

We'll be working with the height data again, so let's filter out the NAs.

```{r}
yrbss = yrbss[!is.na(yrbss$height), ]
```

Next, draw a random sample of these data, with sample size equal to 100.

```{r}
set.seed(1981)
yrbss_samp <- yrbss[sample(x = 1:nrow(yrbss), replace = FALSE, size = 100), ]
```


## Confidence Intervals

Calculate the **sample mean** and the **95% Confidence Interval** for **height** based on your sample. Recall that the **Confidence Interval** is:

$$
\bar{x} \pm 1.96\times SE \\
$$

or, otherwise written as:

$$
\bar{x} \pm 1.96\times \frac{s}{\sqrt{n}}
$$


```{r}

```

## Compare to population mean

Calculate the **population mean** for **height** for the `yrbss` data set. 
Does our **95% confidence interval** calculated in above *include* the true population mean?


```{r}

```



## Z-stat

Using the **population mean** calculated above, and the sample mean calculated even further above, calculate the $z$ statistic associated with your sample mean. Is your sample mean significantly different from the population mean?


```{r}


```


Let's repeat this process 100 times. 

```{r}
p_vals = c()

# start a for loop the does 100 iterations
for(x in 1:100){
  # get a sample of the data of size 100
  yrbss_samp = yrbss[sample(x = 1:nrow(yrbss), replace = FALSE, size = 100), ]
  
  # calculate the new sample mean
  yrbss_samp_mean = mean(yrbss_samp$height)

  # calculate the new SE
  yrbss_samp_SE = sd(yrbss_samp$height)/sqrt(nrow(yrbss_samp))

  # calcualte the Z statistic
  z_stat = (yrbss_samp_mean - yrbss_pop_mean) / yrbss_samp_SE
  
  # calculate the p-value
  p_val = 2 * pnorm(abs(z_stat), lower.tail = FALSE)
  
  # add this p-val to the list
  p_vals = c(p_vals, p_val)
}

```

How many of the sampled data sets are significantly different from the population mean?

```{r}
sum(p_vals < 0.05)
```

That's pretty close to a 5% type I error rate!


# Camparing males and females

Let's now split our sample data by sex, separating male and female measurements.

```{r}
samp_male = subset(yrbss_samp, gender == "male")
samp_female = subset(yrbss_samp, gender == "female")
```

Next, calculate confidence intervals for each of the male and female means.

```{r}

```





