---
title: "Assignment 3 - Working with the Normal Distribution"
output:
  word_document: default
  html_document: default
---

## Assignment Author

<!--- 
Write your name below next to the * sign.
--->

* 

## Working with the YRBSS data set

For this assignment, we are going to use the Youth Risk Behavior Surveillance System (YRBSS) data set.
Let's begin by loading the data into our environment.

```{r}
library(openintro)
data(yrbss)
```

We are going to work with the height data again, so we'll have to perform the `log10` transformation to *normalize* the data.

```{r}
yrbss$log_height = log10(yrbss$height)
```

Let's also remove all of the rows where `height` is NA.

```{r}
yrbss = yrbss[!is.na(yrbss$height), ]
```


Next, let's separate the males and females into two separate data sets.

```{r}
yrbss_female = subset(yrbss, gender == "female")
yrbss_male = subset(yrbss, gender == "male")
```

## Problem 1 

**Calculate the mean and standard deviation values for both the male and female heights.**

```{r}
# put code to calculate mean and sd values here
```


**Describe how these values differ between the males and females. Are these differences expected? Explain your answer.**


## Problem 2

Let's make two histograms of the separated data.

```{r}
library(ggplot2)
ggplot(data = yrbss, aes(x = log_height, fill = gender)) + 
  geom_histogram(position = "dodge")
```

**Make at least three observations regarding what this figure is showing you. What do these combined histograms show you about the differences in heights between males and females?**


## Problem 3

Using **both** an empirical and an analytical approach, calculate the two probabilities below:

* the probability that a randomly chosen male is taller than an average female (i.e., the female mean height). 
* the probability that a randomly chosen female is shorter than an average male (i.e., the male mean height). 


### Code for empirical probability calculations

```{r}
# Put empirical probability calculations here
```


### Code for qnqlytical probability calculations

```{r}
# Put analytical probability calculations here
```

**Explain why there are differences between the probability values calculated using these two methods.**


**Considering the four probability values you just calculated, what does these values say about the differences in height for males versus females? Explain your answer.**






