---
title: "Assignment 6 - Linear Regression"
output:
  word_document: default
---

## Assignment Authors

<!--- 
Write your name below next to the * sign.
--->

* 
* 
* 

#### Grading note

* Problem 1 is worth 10 points
* Problem 2 is worth 10 points
* Problem 3 is worth 5 points

# Overview

For this problem set we will use data on cell survival for fourteen plates exposed to different levels of radiation.
You can acquire these data by using the following command:

```{r}
cell_survival <- read.csv(file = "https://mlammens.github.io/ENS-623-Research-Stats/data/cell_survival.csv")
```

# Problem 1

Use the `lm` function to fit a linear regression model to these data. 
Report the results and plot the data and the best-fit line. (10 pnts)

```{r}
# Put your code in here
```


**Put your written answers here.**


# Problem 2

Use the regression diagnostics we learned about in class to examine these data.
Identify the outlier(s). 
Remove the outlier(s) and run a new `lm`. 
How does the regression slope change if it/they were to be removed from the dataset? 
Be sure to clearly state which data points you are considering "outliers" and why. (10 pnts)

```{r}
# Put your code in here
```


**Put your written answers here.**



# Problem 3

Do these data fit the assumptions of linear regression?
Address each of the "big 3" assumptions separately. (5 pnts)

**Put your written answers here.**