---
title: "Assignment 5 - Two-sample Hypothesis Testing and ANOVA"
output:
  word_document: default
  html_document: default
---

## Assignment Author

<!--- 
Write your name below next to the * sign.
--->

*
* 
* 

# Overview

In this problem set, you are going to work on comparing groups of data, to see if there are significant differences between groups.
We will be using data on Phosphorus levels in the soil at Pace University.

First, let's get the data.


```{r soil_data}
soil_data <- read.csv("https://www.dropbox.com/s/b8vanatf6kejuba/soil_p_data.csv?dl=1")
summary(soil_data)
```

## Visualize the data (5 points)

Use a boxplot to visualize the differences between the three different **vegetation types**. 


```{r}



```

## $t$ test assumptions (5 points)

Based on the information given and this visualization, do you think these data meet the assumptions for a $t$ test? Address each of the three assumptions in your answer.

**Write your answer below this line.**


## $t$ test (10 points)

Regardless of your answer to the question above, assume you can use a $t$ test to compare the different `vegetation_types`. 
Because the $t$ test is only appropriate for comparing at most two groups of data, **you will have to run three different tests to compare each group to each other**. 


```{r}

```

Based on the $t$ test results, can you conclude any of the groups are different from each other?

**Write your answer below this line.**


# ANOVA with R (5 points)

Use either `summary(aov())` or `anova(lm())` to perform an ANOVA for this data set.

**Put your R code below.**

```{r}



```

Do your conclusions about the means of the different vegetation types based on the ANOVA match your conclusions based on the three *t* tests?

