---
title: "Week 9 - t tests and ANOVA - Class Plan"
author: "Matthew Aiello-Lammens"
date: "10/19/2019"
output:
  word_document:
    toc: yes
---

# Class plan

* Demonstration of two-sample *t* test in R
* Demonstration of ANOVA in R

# Demonstration of two-sample *t* test in R

In these demonstrations we will be using the a data set on **aldrin** concentration values at three different depths in the Wolf River.
This case study is further described in the ANOVA lecture.

```{r}
aldrin <- read.csv("https://www.dropbox.com/s/sfstj4p8beajir5/aldrin.csv?dl=1")
summary(aldrin)

```

Let's make a box plot of these data

```{r}
library(ggplot2)

ggplot(data = aldrin, aes(x = depth, y = aldrin)) +
  geom_boxplot()
```

Just looking at the box plot, it does appear as though there are differences in the levels of aldrin at the different depths.

## Compare concentrations of aldrin at different depths with *t* test

There are three possible **pair-wise comparisons** we can do here:

* bottom - middle
* bottom - surface
* middle - surface


We will carry out a 2-sample *t* test for each of these comparisons by following these procedures:

1. Subset the data to include only the two depths of interest
2. Perform a *t* test using the R function `t.test` with the **formulat** notation (i.e., using the ~ sign).

### bottom - middle comparison

Subset our data set

```{r}
aldrin_bot_mid = subset(aldrin, depth %in% c("bottom", "middepth"))
```

Perform *t* test

```{r}
t.test(data = aldrin_bot_mid, aldrin ~ depth)
```

### bottom - surface


```{r}


```


### middle - surface

```{r}


```

# Demonstration of ANOVA in R

## Carry out an ANOVA

ANOVA is a kind of linear model (we'll talk more about this soon), so we use the `lm` function in R first.

```{r}

```

Next, we use a special summary command called `anova` to show the results of our linear model.

```{r}

```

You could do something similar with the `aov` function. 

```{r}


```



In lecture, I mention the possibility of doing multiple comparisons to test for which groups are different from each other.
In practice, we would normally use a method like Tukey's Honset Significant Differences test, which I implement below. Note that I use the `aov` command here, instead of the `anova(lm())` syntax. 

```{r}

```

