---
title: "Week 9 - MON - Hypothesis Testing using the t-test - 2 samples - Class Plan "
author: "Matthew Aiello-Lammens"
date: "11/01/2021"
output: 
  html_document: 
    toc: yes
---

# Learning Objectives

* Practice the basic principles of hypothesis testing
* Become familiar with the **one-sample t-test**
* Identify when to apply the **one-sample t-test** versus the **z-test**
* Become familiar with the **two-sample *t*-test**

# Comparing groups

```{r}
s1 <- rnorm(n = 1000, mean = 2, sd = 0.9)
s2 <- rnorm(n = 1000, mean = 5, sd = 0.9)

samp_data <- data.frame(group = rep(c("s1","s2"), each = 1000), value = c(s1, s2))

library(ggplot2)

ggplot(data = samp_data, aes(x = value)) +
  geom_density(aes(color = group)) +
  xlim(-15, 25) +
  theme_bw()

```


```{r}
s1 <- rnorm(n = 1000, mean = 2, sd = 5)
s2 <- rnorm(n = 1000, mean = 5, sd = 5)

samp_data <- data.frame(group = rep(c("s1","s2"), each = 1000), value = c(s1, s2))

library(ggplot2)

ggplot(data = samp_data, aes(x = value)) +
  geom_density(aes(color = group)) +
  xlim(-15, 25) +
  theme_bw()

```


## Embyonic Stem Cell Research

Read in data. Note use of `read.table` for the .txt file.

```{r}
library(openintro)
data("stem_cell")
```

View data.

```{r}
stem_cell
```

Calculate differences

```{r}
stem_cell$diff <- stem_cell$after - stem_cell$before
```

Calculate summary stats

```{r}
library(dplyr)

stem_cell %>%
  group_by(trmt) %>%
  summarise(mean = mean(diff),
            sd = sd(diff))

```

### Visualize the data

```{r}
library(ggplot2)

ggplot(stem_cell, aes(x = trmt, y = diff)) +
  geom_boxplot() +
  theme_bw()
```

Alternatively

```{r}
ggplot(stem_cell, aes(x = diff, color = trmt)) +
  geom_histogram(position = "dodge", aes(fill = trmt, y = ..density..)) +
  geom_density() +
  theme_bw()
```

### Calculate T-statistic

```{r}
mean_diff <- 3.5 - -4.333333

se_diff <- sqrt((2.764168^2/8) + (5.172040^2/8))

T_diff <- mean_diff / se_diff
```

Now calculate the probability associated with this T-statistic

```{r}
2*pt(-T_diff, df = 8)
```

### Using `t.test`

As with the one-sample t-test we saw on Wednesday, we can do a two-sample t-test using R's built in `t.test` function

In this case, we will be introduced to R's `~` operator.

```{r}
t.test(stem_cell$diff~stem_cell$trmt)
```


