---
title: "Meeting 13 - Distributions of Random Variables in R - Class Plan"
author: "Matthew E. Aiello-Lammens"
output: 
  html_document: 
    toc: yes
---

## Learning Objectives

* Generate a data set that follows a **binomial distribution** using computer simulation methods.
* Calculate analytically and through simulation the probability of observing a specific value for a *random variable*.
* Use the parameters of a probability distribution to calculate the mean and standard deviation of a random variable.



## Generate a binomial distribution through *simulated* coin flips

For this activity, you will be simulating a coin flip several times and recording the number of **heads** you observe.

### Experiment 1 - simulating individual coin flips

For experiment 1, we want to simulate the following:

1. Flip your coin 10 times. Record the number of times a **heads** appeared out of the 10 flips.
2. Repeat step 1 a total of 10 times. By the end of this experiment, you should have 10 sets of 10 flips, with 10 values for the number of heads flipped in each set.

#### Flip a single coin

```{r}
coin_outcomes = c("head", "tail")

sample(coin_outcomes, replace = TRUE, size = 1)
```

#### Flip a coin 10 times

```{r}
sample(coin_outcomes, replace = TRUE, size = 10)

# Save a set of flips
ten_flips = sample(coin_outcomes, replace = TRUE, size = 10)

# Count the heads
sum(ten_flips == "head")
```

#### Do 10 sets of 10 flips

```{r}
heads_count = c()
for(x in 1:10){
  ten_flips = sample(coin_outcomes, replace = TRUE, size = 10)
  ten_flips_heads = sum(ten_flips == "head")
  heads_count = c(heads_count, ten_flips_heads)
}

```



Once you have the number of heads for all 10 sets of 10 flips, calculate the following:

* The average number of heads out of the 10 sets of 10 flips
* The standard deviation for the number of heads out of the 10 sets of 10 flips

```{r}
mean(heads_count)
sd(heads_count)
```


Next, calculate the probability of getting 5 out of 10 heads **BASED ON THE DATA YOU GENERATED**. 
To calculate this, you should divide the number of times you got 5 heads by the total number of sets.

```{r}
sum(heads_count == 5) / length(heads_count)
```


### Exp 1 - Simulations using `rbinom`

For the next step, start RStudio. 
You are going to use RStudio to **simulate** the same experiment you just completed above.

In order to simulate the results of 10 sets of data as we did above, we are going to use the function `rbinom`. 
The inputs to `rbinom` are:

* `n` = the number of *sets* of data to simulate.
* `size` = the *sample size* for each of sets of data. **IMPORTANT** - in the text book, this number is symbolized by *n* and is noted as the number of **independent trials**.
* `prob` = the probability of a *success* for a single trial. In this case, the probability of getting a heads in a single coin flip. **IMPORTANT** - in the text book, this number is symbolized by *p*.

```{r}
coin_flip_heads <- rbinom(n = 10, size = 10, prob = 0.5)
```

The result of running the code above is a vector named `coin_flip_heads` that has a length of 10. 
The numbers represent the number of heads in each of 10 sets of 10 flips.

Use this vector to calculate the average and standard deviation of the number of heads in 10 sets of 10 flips.

```{r}
mean(coin_flip_heads)
sd(coin_flip_heads)
```

Using the RStudio generated data, calculate the probability of getting exactly 5 heads. This calculation is identical to how you calculated the probability above, but the code is a little more complicated.

```{r}
sum(coin_flip_heads == 5) / length(coin_flip_heads)
```


### Exp 1 - Analytical solutions

Flipping a coin a set number of times, as you just did, is an example of a random process whose outcome follows a binomial distribution. 
Because of this, you can calculate the probability of getting exactly 5 heads when you flip a coin 10 times using the analytical solution. 
We will do this in two ways.

First, let's calculate this using the equation for the binomial distribution:

$$
P(X = x)=\binom{n}{x} p^x (1-p)^{n-x}
$$

A few things to recall / know:

* Here *n* is the number of trials, which is 10
* *x* is the number of successes, so here, 5 heads
* *p* is the probability of success, so 0.5
* The function for $\binom{n}{x}$ in R is `choose(n, x)`

```{r}
# Use this space to calculate the probablity of getting 5 heads out of 10 tosses
```


Second, let's use R to do all that work for us. 
Recall that in the lecture, I noted that the function `dbinom` can be used to calculate the exact probability of $X = k$.
Run the code below to see this in action.

```{r}
dbinom(x = 5, size = 10, prob = 0.5)
```

For our last calculations related to Experiment 1, use the equations for the mean value of $X$ and the standard deviation of $X$. 
We will go over this in lecture later, but the equations are:

$$
\text{mean} = n*p
$$

and

$$
\text{standard deviation} = \sqrt{np(1-p)}
$$

***

## Experiment 2

>What is the effect of **EITHER** increasing the number of trials in each set of flips **OR** increasing the number of sets of flips?

For this second experiment, you get to choose whether you want to examine the effects of increasing the number of trials in each set of flips *or* increasing the number of sets.

**Increasing trials** - to increase the number of trials, using the `rbinom` function you would increase the `size` variable. Like wise for the `dbinom` function. 

**Increasing sets** - to increase the number of sets, using the `rbinom` function you would increase the value of `n`. 

**IMPORTANT - CONFUSING SYMBOLOGY** - remember, for the R functions `size` is the same as the symbol $n$ in the equations from the textbook (and above). There is no analog for how R uses `n` in the equations. 

Calculate the following values using the new data you generated by either increasing the number of sets or increasing the number of trials.

* Average (mean) number of heads.
* Standard deviation of the number of heads.
* Probability of 50% of your tosses being heads (i.e., 5 out of 10 heads, or 25 out of 50, etc.).



