---
title: "Week 6 - Class Plan - Confidence Intervals"
author: "Matthew Aiello-Lammens"
date: "9/30/2020"
output:
  html_document:
    toc: yes
---

## Learning Objectives

* Identify the relationships between population statistics and sample statistics
* Calculate probabilities associated with getting certain sets of values for normally distributed data using both empirical approximation and analytical calculations

## Working with the Youth Risk Behavior Surveillance System (YRBSS) data

In this lesson, we're going to explore variation in point estimates and Confidence Intervals.
We'll be working with the Youth Risk
Behavior Surveillance System (YRBSS).
Here, we will start be examining the **full** data set.
(NOTE: We will largely be following content that can be found in the textbook chapter 4 of OpenIntro Statistics, 3rd Edition, by Diez, Barr, and Çetinkaya-Rundel).

### Population parameters

Let's begin by loading the data into our environment.

You may need to first install the `openintro` R package

```{r, eval=FALSE}
install.packages("openintro")
```

After that, we can load in the data.

```{r}
library(openintro)
data(yrbss)
```


As with any data set, let's have a quick look at the data using `summary`.

```{r}
summary(yrbss)
```


For now, let's focus on the variable `height`. 
First, let's calculate the **population mean** for `height`.
Remember, that the population mean is going to be the mean value for the *full* data set.

```{r}
mean(yrbss$height)
```

That's probably not the answer you were expecting. 
There must have been students whose data was collected, but their heights weren't recorded, so a value of *NA* was assigned.
In order to ignore these data, we need to add the argument `na.rm = TRUE` to our `mean` call.

```{r}
mean(yrbss$height, na.rm = TRUE)
```


### Sample parameter estimates

OK, let's assume that we could only collect data from a subset of the high school students whose data are in this data set. 
We can use the data in this subset to **estimate** population level parameters.
This subset would be called our **sample**.

We'll choose a subset at random, but we'll make it so that everyone in the class gets the same random subset.

```{r}
set.seed(1981)
yrbss_samp = yrbss[sample(x = 1:nrow(yrbss), replace = FALSE, size = 10), ]
```

**Challenge:** Breakdown each piece of the command that we are using to get our sample.
To do this, you should make sure to look at the help for `sample`.

Let's calculate the mean height based on our sample.

```{r}
mean(yrbss_samp$height, na.rm = TRUE)
```

How does it compare to the true mean?

Let's try increasing our sample size, and see if that makes a difference?

#### Sample size = 100

```{r}
set.seed(1981)
yrbss_samp = yrbss[sample(x = 1:nrow(yrbss), replace = FALSE, size = 100), ]
mean(yrbss_samp$height, na.rm = TRUE)
```

#### Sample size = 1000

```{r}
set.seed(1981)
yrbss_samp = yrbss[sample(x = 1:nrow(yrbss), replace = FALSE, size = 1000), ]
mean(yrbss_samp$height, na.rm = TRUE)
```


In general, though not always, as you increase sample size, you get closer to the *true* population values for an estimated parameter.


## Working with normally distributed data

Let's use some of the plotting techniques we learned earlier in this course to examine the distribution of the height data.

```{r}
library(ggplot2)

ggplot(data = yrbss, aes(x = height)) + 
  geom_histogram()
```

We could also plot these data with an approximate density curve.

```{r}
ggplot(data = yrbss, aes(x = height, y = ..density..)) + 
  geom_histogram() +
  geom_density()
```

### Data transformations

These data are not quite normally distributed (a little bit right skewed). 
In order to make the data appear normal, we could apply a transformation to each data observation.
The `log10` transformation is fairly standard in statistics.

```{r}
yrbss$log_height = log10(yrbss$height)

ggplot(data = yrbss, aes(x = log_height)) + 
  geom_histogram()
```


And that's about as good as we're going to get!

### Distribution of a sample of the data

Let's have a look at what the distribution of our data looks like if we take only a sample?

```{r}
set.seed(1981)
yrbss_samp = yrbss[sample(x = 1:nrow(yrbss), replace = FALSE, size = 10), ]

ggplot(data = yrbss_samp, aes(x = log_height)) + 
  geom_histogram()
```

**Challenge:** What happens as you increase the sample size?

#### Aside: Simulating normally distributed data

The **Normal Distribution** is defined by two parameters -- the **mean** and the **standard deviation**.
We can *simulate* normally distributed data using the `rnorm` function, with a mean and standard deviation.

Here we are simulating 100 data points that are 'random normally' distributed: 

```{r}
rand_data = rnorm(n = 100, mean = 0, sd = 1)

ggplot(data = NULL, aes(x = rand_data)) +
  geom_histogram()
```


**Challenge:** Change the `mean` value. Describe what happens.

**Challenge:** Change the `sd` value. Describe what happens.

**Challenge:** Change the sample size (`n`) value. Describe what happens.


### Parameterizing a normal with emprical data

We can simulate data that is similar to the YRBSS height data if we parameterize a normal distribution with `mean` and `sd` values (population or sample values) calculated from the data set.

```{r}
yrbss_log_height_mean = mean(yrbss$log_height, na.rm = TRUE)
yrbss_log_height_sd = sd(yrbss$log_height, na.rm = TRUE)
```

Now, draw 10,000 random values from this distribution. Construct a histogram and compare to the histogram above.

```{r}
rand_yrbss_data = rnorm(n = 10000, mean = yrbss_log_height_mean, sd = yrbss_log_height_sd)

ggplot(data = NULL, aes(x = rand_yrbss_data)) +
  geom_histogram()
```

### Probability of certain values of heights

Let's calculate the probability of randomly selecting a study participant who is taller than the mean height. 

**Challenge:** What should this probability be, if the data are perfectly normally distributed?


#### Emperical probability calculations

Here are the steps for an empirical calculation.

* Determine how many values in your data set fit the desired criteria (so here, how many values are greater than the average)
* Divide the number from the first step by the total number of values

```{r}
tot_greater_mean = sum(yrbss$log_height >= yrbss_log_height_mean, na.rm = TRUE)
tot_values = sum(!is.na(yrbss$log_height))

tot_greater_mean / tot_values
```


#### Analytical probability calculations

We can use Rs internal `pnorm` function to calcualte the probability of getting a value greater than the mean.
(NOTE: for *greater* we need to include the `lower.tail = FALSE` argument)

```{r}
pnorm(q = yrbss_log_height_mean, mean = yrbss_log_height_mean, sd = yrbss_log_height_sd, lower.tail = FALSE)
```

#### Example - probability of youth taller than 6 feet

6 feet = 1.8288 meters

```{r}
# log10(1.8288)

# empirical calculation
tot_greater_sixft = sum(yrbss$log_height >= log10(1.8288), na.rm = TRUE)
tot_greater_sixft/tot_values

# analytical calculation
pnorm(q = log10(1.8288), mean = yrbss_log_height_mean, sd = yrbss_log_height_sd, lower.tail = FALSE)

```



