---
title: "MAT 141: Practice with boxplots"
format: 
  docx:
    reference-doc: "/Users/maiellolammens/bw-reference.docx"
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ex <- data.frame(mohonk = c(0, 1, 4, 4, 5, 6, 6, 7, 8, 8, 10, 11),
                 teatown = c(2, 3, 6, 6, 7, 8, 8, 9, 10, 10, 12, 23))
```

## Names:

&nbsp;

## Introduction

Tracey was interested to see if there was a difference in the number of white-footed mice (WFM) at two different preserves, Mohonk Preserve and Teatown Lake Reservation.
She placed 12 camera traps out in each preserve for 2 weeks. 
She recorded the number of white-footed mouse (WFM) pictures taken by each camera during those two weeks,
and found the following counts at the two preserves.

### Mohonk Preserve WFM Counts

```{r, echo=FALSE, message=FALSE, warning=FALSE}
cat(sample(ex$mohonk), sep = "   ")
```

&nbsp;

### Teatown Lake Reservation WFM Counts

```{r, echo=FALSE}
cat(sample(ex$teatown), sep = "   ")
```

&nbsp;

For each set of WFM counts, calculate the summary statistics required to construct a box plot :

* Median
* 25% and 75% quartiles
* IQR

and use these values to construct **two** box plots side-by-side below.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
ggplot(data = ex, aes(y = teatown)) +
  ylim(c(0,25)) +
  geom_blank() + 
  theme(axis.title.y=element_blank())

```

