---
title: "Week 7 - Class Plan - Hypothesis Testing"
author: "Matthew Aiello-Lammens"
date: "10/07/2020"
output:
  word_document:
    toc: yes
  html_document:
    toc: yes
---

# Lecture on Hypothesis Testing

First half of class will be dedicated to hypothesis testing introduction lecture.

# Worked example

Second half of lecture will be dedicated to a worked example of a hypothesis test.

## *More* work with the Youth Risk Behavior Surveillance System (YRBSS) data

This week, we are again going to use the Youth Risk Behavior Surveillance System (YRBSS) data set.
Let's begin by loading the data into our environment.

```{r}
library(openintro)
data(yrbss)
```

Next, draw a random sample of these data, with sample size equal to 100.

```{r}
set.seed(1981)
yrbss_samp <- yrbss[sample(x = 1:nrow(yrbss), replace = FALSE, size = 100), ]
```


## Confidence Intervals

Calculate the **sample mean** and the **95% Confidence Interval** for **height** based on your sample. Recall that the **Confidence Interval** is:

$$
\bar{x} \pm 1.96\times SE \\
$$

or, otherwise written as:

$$
\bar{x} \pm 1.96\times \frac{s}{\sqrt{n}}
$$


```{r}
mean(yrbss_samp$height, na.rm = TRUE)

mean(yrbss_samp$height, na.rm = TRUE) + 1.96*(sd(yrbss_samp$height, na.rm = TRUE)/sqrt(100))
mean(yrbss_samp$height, na.rm = TRUE) - 1.96*(sd(yrbss_samp$height, na.rm = TRUE)/sqrt(100))

```

## Compare to population mean

Calculate the **population mean** for **height** for the `yrbss` data set. 
Does our **95% confidence interval** calculated in above *include* the true population mean?


```{r}
mean(yrbss$height, na.rm = TRUE)
```

> In my case, my 95% **DOES** include the population mean. For the most part, we should expect this to be the case, though we could get a set of data in our sample that happens to be extreme, and therefore, the CI does not include the mean.


## Z-stat

Using the **population mean** calculated above, and the sample mean calculated even further above, calculate the $z$ statistic associated with your sample mean. Is your sample mean significantly different from the population mean?


```{r}
# Get the sample mean (again)
yrbss_samp_mean <- mean(yrbss_samp$height, na.rm = TRUE)

# Get the population mean (again)
yrbss_pop_mean <- mean(yrbss$height, na.rm = TRUE)

# Get the Standard Error of the Mean for the sample (again)
yrbss_samp_SE <- sd(yrbss_samp$height, na.rm = TRUE)/sqrt(100)

## Calcualte the Z statistic
z_stat <- (yrbss_samp_mean - yrbss_pop_mean) / yrbss_samp_SE
print("my z_stat value")
z_stat

## Calculate the probability of this Z-stat - assume a **TWO TAILED** test
## Below, I'm using the absolute value of the z_stat value, then
## looking at the upper tail only.
print("probability of my z_stat value, assuming two-tailed test")
2 * pnorm(abs(z_stat), lower.tail = FALSE)

```


> Looking at the probability value for my Z-stat value, assuming an $\alpha = 0.05$ significane value, we DO NOT reject the null hypothesis. Thus, we support the hypothesis that my sample mean is NOT significantly different than the population mean.



