---
title: "Week 13 - Tuesday - more ANOVA - Class Plan"
author: "Matthew Aiello-Lammens"
output:
  html_document:
    toc: yes
    df_print: paged
---

# Class plan

* Review of ANOVA
* Demonstration of ANOVA in R
* Begin assignment 5


# Demonstration of ANOVA in R

This is a review of materials from our Wednesday 11/13 class. 
However, this time we are going to work through the ANOVA "by hand".

We are going to work with the data set on aldrin levels at differnet water depths of the Wolf River.

```{r}
#aldrin <- read.csv("https://www.dropbox.com/s/sfstj4p8beajir5/aldrin.csv?dl=1")
aldrin <- read.csv("https://bit.ly/2DXjm59")
summary(aldrin)

```

Let's make a box plot of these data

```{r}
library(ggplot2)

ggplot(data = aldrin, aes(x = depth, y = aldrin)) +
  geom_boxplot()
```

As before, we can say that just looking at the box polt, it does appear as though there are differences in the levels of aldrin at the different depths.

We can also visualize these data as a scatter plot.

```{r}
ggplot(data = aldrin, aes(x = depth, y = aldrin, color = depth)) +
  geom_point(position = "jitter") +
  theme_bw()
```


## Carry out an ANOVA *by hand*

### Let's start by calculating the sums of squares total

$$
SST = \sum_{i=1}^n (x_i -\bar{x})^2
$$
, where *n* is the total number of observations.

```{r}
# Calculate the grand mean
grand_mean <- mean(aldrin$aldrin)

# Calcualte the differences between the individual observations and the grand mean
diffs <- aldrin$aldrin - grand_mean

# Calculate the squared differences (or squared deviations of the observations from the mean)
sqr_diffs <- diffs^2

# Calculate the sum of squared differences
(ss_total <- sum(sqr_diffs))

```


### Next calculate the sums of squares groups

$$
SSG = \sum_{i=1}^k n_i(\bar{x}_i -\bar{x})^2
$$
, where *k* is the number of groups.

```{r}
# Calculate the grand mean (again)
grand_mean <- mean(aldrin$aldrin)

# Make subsets of the data for each of the depths
unique(aldrin$depth)

aldrin_bottom <- subset(aldrin, depth == "bottom")
aldrin_middepth <- subset(aldrin, depth == "middepth")
aldrin_surface <- subset(aldrin, depth == "surface")

# Calculate the group means
bottom_mean <- mean(aldrin_bottom$aldrin)
middepth_mean <- mean(aldrin_middepth$aldrin)
surface_mean <- mean(aldrin_surface$aldrin)

# Calculate the sums of squared groups
ss_group <-
  sum(10*(bottom_mean - grand_mean)^2 +
      10*(middepth_mean - grand_mean)^2 +
      10*(surface_mean - grand_mean)^2)

ss_group

```


### Next calculate the sums of squares error (or residual)

As we saw in class two weeks ago, there are two ways to do this. First is to acknowledge that:

$$
SST = SSG + SSE
$$

So one way to calculate SSE is the following:

```{r}
ss_error_1 <- ss_total - ss_group
ss_error_1
```

Another way is to calculate the sums of squares error directly, following the equation:

$$
SSE = \sum_{i=1}^n (x_i - \bar{x}_{i,k})^2 
$$
, where $\bar{x}_{i,k}$ is the mean for the group $k$ that a particular $x_i$ belongs too.

```{r}
ss_error_2 <-
  sum(sum((aldrin_bottom$aldrin - bottom_mean)^2),
      sum((aldrin_middepth$aldrin - middepth_mean)^2),
      sum((aldrin_surface$aldrin - surface_mean)^2))
ss_error_2
```

### Now calculate Mean Sums of Squares

Recall that the Mean Sums of Squares Groups is:

$$
MSG = SSG / df_{group}
$$

, where $df_{group}$ is the number of groups minus 1, i.e. $k-1$.

And the Mean Sums of Squares Error is:

$$
MSE = SSE / df_{error}
$$

, where $df_{error}$ is the sum of the number of observations is each group minus 1.

So to calculate the MSG and MSE, we do

```{r}
df_group <- (3-1)
MSG <- ss_group / df_group
MSG
```

and

```{r}
df_error <- sum((nrow(aldrin_bottom) - 1),
                        (nrow(aldrin_middepth) - 1),
                        (nrow(aldrin_surface) -1))
MSE <- ss_error_2 / df_error

MSE
```


### Calculate the F-stat for this data set

```{r}
f_stat <- MSG/MSE
```

### Finally, calcualte the probability of this particular F value

```{r}
pf(q = f_stat, df1 = df_group, df2 = df_error, lower.tail = FALSE)
```



## Carry out an ANOVA using Rs internal functions

ANOVA is a kind of linear model (we'll talk more about this next week), so we use the `lm` function in R first.

```{r}
aldrin_lm <- lm(data = aldrin, aldrin~depth)
```

Next, we use a special summary command called `anova` to show the results of our linear model.

```{r}
anova(aldrin_lm)
```

You could do something similar with the `aov` function. 

```{r}
aldrin_aov <- aov(data = aldrin, aldrin~depth)
summary(aldrin_aov)
```



In the past, we dicussed the possibility of doing multiple comparisons to test for which groups are different from each other.
In practice, we would normally use a method like Tukey's Honset Significant Differences test, which I implement below. Note that I use the `aov` command here, instead of the `anova(lm())` syntax. 

```{r}
TukeyHSD(aov(data = aldrin, aldrin~depth))
```

